---
title: "diamonds_dataset_exploration"
author: "aditya"
date: "Today's Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Introduction

In this tutorial, we will be covering EDA. Specifically, we will use data visualization and transformation to explore the `diamonds` dataset.


1. What type of variation occurs within my variables?
2. What type of covariation occurs between my variables?


```{r}
glimpse(diamonds)
```

## Variation

- Categorical Variables

```{r bar_plot, fig.cap="Bar Chart Looking at Diamond Cut"}
ggplot(diamonds) +
  geom_bar(mapping=aes(x=cut))
```

```{r cut_count}

diamonds %>% count(cut)

```

- Continuous Variables

```{r hist_carat}

ggplot(diamonds) +
  geom_histogram(aes(x=carat), binwidth=0.5)

```
## Data Cleaning

```{r small_dataset_creation}

small <- diamonds %>% filter(carat < 3)

```


```{r small_hist}

ggplot(small, aes(x=carat)) +
  geom_histogram(binwidth=0.1)

```
## Typical Values

1. Which values are the most common? Why?
2. Which values are rare? Why? Does that match your expectations?
3. Can you see any unusual patterns? What might explain them?

```{r}

ggplot(small, aes(x=carat)) + geom_histogram(binwidth=0.01)

```
**Questions that need to be answered**
- Why are more diamonds at whole carats?
- Why are there more diamonds that are more to the right of the common carats?


## Covariation


### Categorical vs Continuous

```{r}

ggplot(diamonds) + geom_bar(aes(x=cut))

```

### Categorical vs Categorical

```{r}

ggplot(diamonds, aes(x=price)) + 
  geom_freqpoly(aes(x=price)) +
  geom_freqpoly(aes(color=cut), binwidth=500)

```
- Boxplot

### Continuous vs Continuous

```{r}

ggplot(diamonds, aes(x=cut, y=price)) + 
  geom_boxplot()

```


```{r}

diamonds %>% count(color, cut) %>% ggplot(aes(x=color, y=cut)) +
  geom_tile(aes(fill=n))

```


```{r}

ggplot(diamonds) + 
  geom_point(aes(x=carat, y=price), alpha=1/100)

```


```{r}

ggplot(diamonds) + geom_hex(aes(x=carat, y=price))

```


```{r}

ggplot(diamonds, aes(x=carat, y=price)) + 
  geom_boxplot(aes(group=cut_width(carat, 0.1)))

```

## Patterns and Models


```{r}

library(modelr)

```


```{r}

mod <- lm(log(price) ~ log(carat), data=diamonds)

```


```{r summary_mod}

summary(mod)

```

```{r lm_resid}

diamonds2 <- diamonds %>% add_residuals(mod) %>% mutate(resid=exp(resid))

```

```{r}

glimpse(diamonds2)

```



```{r}

ggplot(diamonds2) + geom_point(aes(x=carat, y=resid), alpha=1/50)

```

